---
title: "Police Violence Project"
subtitle: "An investigation into inequities"
author: "Justin Pimentel and Joseph Snitzer"
output:
  tufte::tufte_html: default
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(leaflet)
library(sf)
library(sp)
library(geofacet)
library(lubridate)
# remotes::install_github("wilkelab/ggtext")
library(ggtext)
mapthm <-  theme(axis.title.x=element_blank(), axis.text.x=element_blank(),
                 axis.ticks.x=element_blank(), axis.title.y=element_blank(),
                 axis.text.y=element_blank(),axis.ticks.y=element_blank(),
                 panel.grid.major = element_blank(),panel.border = element_blank(),
                 plot.background = element_rect(fill = "#FFFFF8", color = NA))
```

# Time series by state

> "bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla "
> --- [bla  bla ](https://google.com)

bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla ^[bla bla bla ].

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
### Here we're importing and merging our data

data <- read.csv("data/MPV.csv", header=T)
data$State <- as.character(data$State)
df <- read.csv("data/Addresses.csv", header=F)
colnames(df) <- c("ID","x","y")
data <- merge(data, df, by="ID")
data_va <- data[which(data$State == "VA"),]
data_mn <- data[which((data$State %in% c("HI","AK"))==F),]



### Here we're adding our data to our states shapefile for results

States <- read_sf("data/States/cb_2018_us_state_20m.shp")
States <- States[which((States$NAME %in% c("Puerto Rico","Hawaii","Alaska")==F)),]
data_mn[which(data_mn$BodyCamera %in% c("Bystander Video","no","No","Surveillance Video")),"BodyCamera"] <- "No"



### Here we're extracting the race murder data, by state, by year
state_cases <- data.frame(data_mn %>% group_by(State) %>% summarise(Killings = length(ID)))
States2 <- merge(States, state_cases, by.x="STUSPS", by.y="State")
data_mn$Date <- as.Date(data_mn$Date,format="%m/%d/%y")
data$Date <- as.Date(data$Date,format="%m/%d/%y")
data$Year <- format(data$Date,"%Y")
data$BlackDeath <- 0
data[which(data$Race=="Black"),"BlackDeath"] <- 1
data$WhiteDeath <- 0
data[which(data$Race=="White"),"WhiteDeath"] <- 1
graph_data <- data.frame(data %>% 
                         group_by(State, Year) %>% 
                         summarise(Black = sum(BlackDeath), White = sum(WhiteDeath)))
AvgDeaths <- data.frame(graph_data %>% group_by(Year) %>% 
                         summarise(Black = sum(Black), White = sum(White)))
AvgDeaths$TotalBlack <- NA
AvgDeaths$TotalWhite <- NA



### Standardizing the results by race

for(i in 2013:2019)
{
  if(i==2019)
  {
    pops <- read.csv(paste("data/RacePop/ACSDT1Y",2018,
                         ".B02001_data_with_overlays_2020-05-27T221318.csv",sep=""))
  }
  else
  {
    pops <- read.csv(paste("data/RacePop/ACSDT1Y",i,
                           ".B02001_data_with_overlays_2020-05-27T221318.csv",sep=""))
  }
  pops <- pops[2:nrow(pops),c("NAME","B02001_001E","B02001_002E","B02001_003E")]
  colnames(pops) <- c("StateName","Total","White","Black")
  pops$White <- as.numeric(as.character(pops$White))
  pops$Black <- as.numeric(as.character(pops$Black))
  pops$State <- state.abb[match(pops$State,state.name)]
  pops[which(pops$StateName=="District of Columbia"),"State"] <- "DC"
  tempdf <- graph_data[which(graph_data$Year==i),]
  tempdf <- merge(tempdf, pops, by="State")
  tempdf$BlackNew <- tempdf$Black.x/tempdf$Black.y*1000000
  tempdf$WhiteNew <- tempdf$White.x/tempdf$White.y*1000000
  graph_data[which(graph_data$Year==i),"Black"] <- tempdf$BlackNew
  graph_data[which(graph_data$Year==i),"White"] <- tempdf$WhiteNew
  AvgDeaths[which(AvgDeaths$Year==i),"TotalBlack"] <- 
    sum(pops[which(pops$StateName!="Puerto Rico"),"Black"])
  AvgDeaths[which(AvgDeaths$Year==i),"TotalWhite"] <- 
    sum(pops[which(pops$StateName!="Puerto Rico"),"White"])
}

missing <- data.frame(cbind("State" = rep(unique(graph_data$State),each=length(2013:2019)),
      "Year" = rep(2013:2019,length(unique(graph_data$State)))))
graph_data <- merge(missing, graph_data, by=c("State","Year"), all=T)
graph_data[which(is.na(graph_data$Black)),c("Black","White")] <- 0

### Graphing the results

AvgDeaths$BlackAvg <- AvgDeaths$Black/AvgDeaths$TotalBlack*1000000
AvgDeaths$WhiteAvg <- AvgDeaths$White/AvgDeaths$TotalWhite*1000000
AvgDeaths <- AvgDeaths[,c("Year","BlackAvg","WhiteAvg")]
graph_data <- merge(graph_data, AvgDeaths, by="Year")

region <- cbind("State"=state.abb, "Region"=state.division)
region <- rbind(region, c("DC","3"))
graph_data <- merge(graph_data, region, by="State")

ggplot(graph_data) + 
  geom_rect(aes(fill = Region), xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, alpha = 0.075) +
  geom_line(aes(x=Year,y=BlackAvg,group=State), color="#e63946", alpha=0.8, linetype = "dashed") +
  geom_line(aes(x=Year,y=WhiteAvg,group=State), color="#457b9d", alpha=0.8, linetype = "dashed") +
  geom_line(aes(x=Year,y=Black,group=State), color="#e63946") +
  geom_line(aes(x=Year,y=White,group=State), color="#457b9d") +
  theme_minimal()+
  facet_geo(~State)+
  theme(plot.title=element_text(face="bold",size=rel(1.2)),
        plot.caption=element_text(hjust=0)) + 
  theme(axis.text.x = element_blank(),
        axis.text.y = element_text(size=6),
        plot.title = element_markdown(lineheight = 1.1, hjust = 0.5),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        plot.background = element_rect(fill = "#FFFFF8", color = NA)) +
  labs(x="",y="",
       title = "**Murders by Police Officers Per Million, by Race, 2013-2019**  
    <span style='font-size:10pt'>
    <span style='color:#457b9d;'>Blue</span>
    represents white murders, and 
    <span style='color:#e63946;'>red</span>
    represents black murders<br>
    Dotted lines indicate national averages, and colors indicate regions") +
  scale_fill_manual(values = c("#ffadad","#ffd6a5","#fdffb6",
                               "#caffbf","#9bf6ff","#a0c4ff",
                               "#bdb2ff","#ffc6ff","#e5e5e5"))


```

# Murder Locations

bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla bla ^[bla bla bla ].

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}

 ggplot() + geom_sf(States, mapping=aes()) + 
   mapthm + 
   geom_point(data_mn[which(data_mn$BodyCamera %in% c("No")),], 
              mapping = aes(x=x, y=y), size=0.5, color="red") + 
   geom_point(data_mn[which(data_mn$BodyCamera %in% c("Yes")),], 
              mapping = aes(x=x, y=y), size=1, color="blue")

```


```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
